---
title: "Panel Analysis"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
output: html_notebook
---

# Preparation
## Get the dataset
```{r}
df <- read.csv("DATA/merge2.csv")
View(df)
df <- df[, -1]
df
View(df)
library(imputeTS)
# df <- imputeTS::na_ma(df, k = 4, weighting = "simple")
# View(df)
na_interpolation(df, option = "linear")
View(df)

# determine if there's na
sum(is.na(df))


# Importing the large dataset of all voila

file <- df

summary(file)
colnames(file)[8] <- "raw_death_rates"
colnames(file)[8]

# Change factory_density's data type from character to numeric
file$factory_density <- as.numeric(as.character(file$factory_density))
```

## Get the packages
```{r}
library(tidyverse) # Modern data science library
library(plm) # Panel data analysis library
library(car) # Companion to applied regression
library(gplots) # Various programing tools for plotting data
library(tseries) # For timeseries analysis
library(lmtest) # For hetoroskedasticity analysis
library(dplyr)
library(showtext)

showtext_auto(enable = TRUE)
```

## Get ready for plm
```{r}
# Check columns' data type
file %>% summarise_all(funs(class))

# An object of class 'pdata.frame' is a data.frame with an index attribute that describes its individual and time dimensions.
pdf <- pdata.frame(file, c("county", "year"))

# View pdf
View(pdf)
```

# Analysis
## The model of all

This is the model that includes all variables:
```
model<- (raw_death_rates ~ waste_recycled_muni + bulk_waste_recycled + rate_material_recov + locomobiles_density + automobiles_density + factory_density + doctor + ozone + revenue + sex + deaths + mean_age)
```

This is the one we'll be using
```{r}
model <- (raw_death_rates ~ waste_recycled_muni + locomobiles_density + automobiles_density + factory_density + urban_density + ozone + revenue)
```

## The 2 "Effects" Model
### Random Effects Model
```{r}
random <- plm(model, data = pdf, model = "random")
summary(random)
```

### Fixed Effects Model
```{r}
fixed <- plm(model, data = pdf, model = "within")
summary(fixed)
```

### Compare the 2 Effects Model
H0) = Random Effects Model is efficient
H1) = Fixed Effects Model is efficient
  if a > 0.05:
    We should use the random effects model
    
```{r}
phtest(fixed, random)
```

## Regression Diagnostics
### Time-fixed effects
```{r}
fixed.time <- plm(raw_death_rates ~ ozone + factor(year) + waste_recycled_muni + locomobiles_density + automobiles_density + factory_density + urban_density + revenue, data = pdf, model = "within")
summary(fixed.time)
```

### Testing Time-fixed effects model. The null hypothesis is that no time-fixed effects are needed
```{r}
pFtest(fixed.time, fixed)
plmtest(fixed, c("time"), type = ("bp"))
```

### If the p value < 0.05 then use time-fixed effects.
### In this example, no need to use time-fixed effects.

## Random effects vs Pooled OLS
```{r}
pool <- plm(raw_death_rates ~ ozone, data = pdf, model = "pooling")
summary(pool)
```


## Breusch-Pagan Lagrange Multiplier for random effects. Null is no panel effect (i.e. OLS better).
```{r}
plmtest(pool, type = c("bp"))
```


## Cross-sectional dependence testing
```{r}
fixed <- plm(raw_death_rates ~ ozone, data = pdf, model = "within")
pcdtest(fixed, c("lm"))
pcdtest(fixed, c("cd"))
```

## Serial Correlation Testing
```{r}
pbgtest(fixed)
```


## Unit Roots / Stationary Testing
### H0) The null hypothesis is that the series has a unit root (i.e. non-stationary)
```{r}
adf.test(pdf$raw_death_rates, k = 2)
```


## Heteroskedasticity testing
If hetersokedasticity is detected we need to use a robust covariance matrix (Sandwich estimator) to account for it
```{r}
bptest(raw_death_rates ~ ozone + factor(county), data = pdf, studentize = F)
```

## Controlling for heteroskedasticity: Random effects
### Original Coefficients
```{r}
coeftest(random)
```

### Heteroskedasticity consistent coefficients
```{r}
coeftest(random, vcovHC)
```

### Heteroskedasticity consistent coefficients, type 3
```{r}
coeftest(random, vcovHC(random, type = "HC3"))
```

### The following shows the HC standard errors of the coefficients
```{r}
t(sapply(c("HC0", "HC1", "HC2", "HC3", "HC4"), function(x) sqrt(diag(vcovHC(random, type = x)))))
```


## Controlling for heteroskedasticity: Fixed effects
### Original coefficients
```{r}
coeftest(fixed)
```
### Heteroskedasticity consistent coefficients
```{r}
coeftest(fixed, vcovHC)
```
### Heteroskedasticity consistent coefficients (Arellano)
```{r}
coeftest(fixed, vcovHC(fixed, method = "arellano"))
```
### Heteroskedasticity consistent coefficients, type 3
```{r}
coeftest(fixed, vcovHC(fixed, type = "HC3"))
```
### The following shows the HC standard errors of the coefficients
```{r}
t(sapply(c("HC0", "HC1", "HC2", "HC3", "HC4"), function(x) sqrt(diag(vcovHC(fixed, type = x)))))
```


